---
title: "Project 2: STAT302project2 Tutorial"
author: "Wuwei Zhang"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{STAT302project2 Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(STAT302project2)
data("my_gapminder")
```

## Introduction

STAT302project2 package contains four functions that are able to 

* perform one sample t-test, 
* fit a linear model, 
* predict output class using k-nearest neighbors cross-validation, 
* predict output class using random forest cross-validation.

In this tutorial, I will explain how to use these functions with `my_penguins` data and `my_gapminder` data.

Install STAT302project2 package using:
```{r, eval = FALSE}
devtools::install_github("wuweizhang0723/STAT302project2")
```

## Tutorial for `my_t.test`

Use the `lifeExp` data from `my_gapminder`.
Demonstrate a test of the hypothesis
  \begin{align}
  H_0: \mu &= 60,\\
  H_a: \mu &\neq 60.
  \end{align}
```{r}
my_t.test(my_gapminder$lifeExp, "two.sided", 60)
```




